<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Introduction}
-->
```{r include=FALSE}
options(markdown.HTML.header = system.file('misc', 'vignette.css', package='knitr'))
```

# sim_sample

After the data generation you may want to draw a sample from the population. If you should be interested in keeping information about your population, see the vignette 'addingComputation'. Use the function `sim_sample()` to add a sampling component to your `sim_setup`. In general you can provide an own sampling function. Two predefined functions are available:

- `sample_number` - wrapper around `dplyr::sample_n`
- `sample_fraction` - wrapper around `dplyr::sample_frac`

```{r}
library(saeSim)
base_id(3, 4) %>% sim_gen_fe() %>% sim_sample(sample_number(1L))
base_id(3, 4) %>% sim_gen_fe() %>% sim_sample(sample_number(1L, groupVars = "idD"))
```

```{r eval=FALSE}
# simple random sampling:
sim_base_lm() %>% sim_sample(sample_number(size = 10L))
sim_base_lm() %>% sim_sample(sample_fraction(size = 0.05))
# srs in each domain/cluster
sim_base_lm() %>% sim_sample(sample_number(size = 10L, goupVars = "idD"))
sim_base_lm() %>% sim_sample(sample_fraction(size = 0.05, groupVars = "idD"))
```